formula1_start_vs_finish.Rmd

### 2024-03-13 inst/formula1_start_vs_finish.Rmd
bugfixes

```{r definemakeH3}
`makeH3` <- function(i,lambda){
  suppressWarnings(a <- setNames(as.numeric(finish[,1]),rownames(finish)))
  suppressWarnings(b <- setNames(as.numeric(grid  [,1]),rownames(finish)))
  wanted <- !is.na(b)
  a <- a[wanted]
  b <- b[wanted]
  a[is.na(a)] <- 0 # did not finish
  H <- ordervec2supp(a)
  for(i in seq_along(b)){H[names(b)[i]] <- as.weight(lambda^(i-1))}
  return(H)
}

likelam <- function(lambda){
  H <- hyper3()
  for(i in seq_len(ncol(finish))){H <- H + makeH3(1,lambda=lambda)}
  return(H)
}
```

```{r evaluatelamrange,cache=TRUE}

f <- function(lambda){
  H <- likelam(lambda)
  loglik(maxp(H,n=4),H)
}
f(0.9)
f(0.901)
```

```{r evalam,cache=TRUE}
lam <- seq(from=0.9,to=1.05,len=5)
like <- sapply(lam,f)
```

```{r plotlikelam}
plot(lam,like,type='b')
```

>>>>

```{r definemakeH3}
replacenonzerowithorder <- function(x){
  nonzero <- x!=0
  x[nonzero] <- order(x[nonzero])
  return(x)
}

order_non_NA_NZ <- function(x){ # leaves NA and 0 entries unchanged; replaces other elements with their order
  wanted <- !is.na(x)
  x[wanted] <- replacenonzerowithorder(x[wanted])
  return(x)
}

topn <- 15
finish <- finish[seq_len(topn),]
options(use_alabama=TRUE)

`makeH3` <- function(i,lambda,topn){
  suppressWarnings(a <- setNames(as.numeric(finish[,i]),rownames(finish)))
  suppressWarnings(b <- setNames(as.numeric(grid  [,i]),rownames(finish)))
  wanted <- !is.na(b)
  a <- a[wanted]
  b <- b[wanted]
  a[is.na(a)] <- 0 # did not finish
  a <- order_non_NA_NZ(a)
  a <- a[seq_len(topn)]
  H <- ordervec2supp(a)
  for(j in seq_along(b)){
    if(!is.na(names(b)[j])){H[names(b)[j]] <- as.weight(lambda^b[j])}
  }
  return(H)
}

likelam <- function(lambda){
  H <- hyper3()
  for(i in seq_len(ncol(finish))){H <- H + makeH3(i,lambda=lambda,topn=topn)}
  return(H)
}
```

```{r evaluatelambbda,cache=TRUE}
f <- function(lambda){
  H <- likelam(lambda)
  m <- maxp(H,n=1)
  return(c(loglik(m,H),lambda,m))
}
```

```{r evalam,cache=TRUE}
lam <- seq(from=0.85, to=1.05, len=15)
bigmat <- sapply(lam,f)
bigmat
```

```{r plotlikelambda}
if(FALSE){plot(lam,like-max(like),type='b')
abline(h=-2)
abline(v=1)
}
```
